---
title: "Mussel Survival 10-12/2020"
author: Katrin Schertenleib
output: github_notebook
---
date when file was first created: Wed Feb 24 15:47:20 2021
update: Wed Nov  2 12:56:13 2022

Aims: 
1. Investigate if treatments had an effect on mussel survival
2. Did deaths occur equally spread across time?

# 1. Housekeeping
libraries

```{r setup}
#library(stringr)
#library(installr) #to always use the latest version
#updateR()
#library(ggplot2) #to make ggplot graphs
library(car) #to run Anova() with different type sums of squares
library(tidyverse) #includes ggplot, dplyr and plyr
library(Hmisc) #to calculate stats for mean error plots in ggplot
```

# 2. Reading in the data
```{r}
Mussel_Survival <- read.csv("2020-12-17_Mussel-Survival.csv")
```

# 3. Data exploration, investigation of total deaths, or survivors, respectively
```{r}
str(Mussel_Survival)
table(Mussel_Survival$Assemblage)

#limit data set to rows containing mussels
MSur.dat <- subset(Mussel_Survival, Mussel_Survival$Assemblage=="KM" | Mussel_Survival$Assemblage=="M")

str(MSur.dat)

table(MSur.dat$Treatment)

hist(MSur.dat$M_dead_AM) #M_dead_AM: mussels dead after manipulation started
boxplot(MSur.dat$M_dead_AM ~ MSur.dat$Treatment)
tapply(MSur.dat$M_dead_AM, MSur.dat$Treatment, summary)
```

percentage death
```{r}
MSur.dat$PD <- (MSur.dat$M_dead_AM/MSur.dat$M_manip)*100
boxplot(MSur.dat$PD ~ MSur.dat$Treatment)
```
looks like 1) there are more deaths in M treatments than in KM; 2) temperature had a negative effect on survival, i.e. the warmer, the more deaths; 3) increased pCO2 didn't have an effect, or a slight positive one (less deaths compared to same temperature level at ambient pCO2)


## 3.1 Exploratory plotting
Percentage death
```{r}
#for better facet labeling I need to rename the levels of Assemblage:
MSur.dat_mod <- MSur.dat %>%    #requires tidyverse
  # Rename levels of Assemblage
  mutate(Assemblage = dplyr::recode(Assemblage, "KM" = "Mussels and algae (MA)", "M" = "Mussels only (M)"),
         Treatment = dplyr::recode(Treatment, "KM_Ta_Ca"="MA_Ta_Ca", "KM_Ta_C+"="MA_Ta_C+", "KM_T+_Ca"="MA_T+_Ca", "KM_T+_C+"="MA_T+_C+", "KM_T++_Ca"="MA_T++_Ca", "KM_T++_C+"="MA_T++_C+"))

#to place Mussels only first in the order of the Assemblage levels
MSur.dat_mod$Assemblage <- factor(MSur.dat_mod$Assemblage, levels = c("Mussels only (M)", "Mussels and algae (MA)")) #specify factor levels in the order you want

#to order the Temperature levels in an increasing way
MSur.dat_mod$Temperature <- factor(MSur.dat_mod$Temperature, levels = c("T++", "T+", "Ta")) #specify factor levels in the order you want

#to order the combined Treatment levels in an increasing way
MSur.dat_mod$Treatment <- factor(MSur.dat_mod$Treatment, levels = c("MA_Ta_Ca", "MA_Ta_C+", "MA_T+_Ca", "MA_T+_C+", "MA_T++_Ca", "MA_T++_C+", "M_Ta_Ca", "M_Ta_C+", "M_T+_Ca", "M_T+_C+", "M_T++_Ca", "M_T++_C+")) #specify factor levels in the order you want


FigPD<-ggplot(MSur.dat_mod, aes(x=Treatment, y=PD, colour=Temperature, fill=pCO2)) # colour needs to be character
FigPD+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  geom_boxplot(lwd=0.8)+ #plots the data 
  scale_colour_manual(values = c("skyblue", "royalblue", "navy"))+
  scale_fill_manual(values = c("grey80", "white"))+  #, labels = expression("...", "...")
  labs(x="Experimental treatments", y="Total mussel deaths [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        #axis.ticks = element_line(size = 0.8),
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=10, angle=90, hjust=0, vjust=0.25), #size of the numbers along the axes; hjust               and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        ) 
```
```{r}
#ggsave(path="Plots", filename="2022-11-02_MusselTotalDeathsPercentage.png", width=8, height=5)
```



Slightly changed graph for presentations (instead of papers): individual labels of experimental treatments removed
```{r}
#for better facet labeling I need to rename the levels of Assemblage:
MSur.dat_mod2 <- MSur.dat %>%    #requires tidyverse
  # Rename levels of Assemblage
  mutate(Assemblage = dplyr::recode(Assemblage, "KM" = "Mussels and algae", "M" = "Mussels only"), # (MA)  (M)
         pCO2 = dplyr::recode(pCO2, "C+" = "650 ppm", "Ca" = "450 ppm (ambient)"),
         Temperature = dplyr::recode(Temperature, "T++" = "ambient +2°C", "T+" = "ambient +1°C", "Ta" = "ambient (9 - 12°C)"),
         Treatment = dplyr::recode(Treatment, "KM_Ta_Ca"="MA_Ta_Ca", "KM_Ta_C+"="MA_Ta_C+", "KM_T+_Ca"="MA_T+_Ca", "KM_T+_C+"="MA_T+_C+", "KM_T++_Ca"="MA_T++_Ca", "KM_T++_C+"="MA_T++_C+"))

#to place Mussels only first in the order of the Assemblage levels
MSur.dat_mod2$Assemblage <- factor(MSur.dat_mod2$Assemblage, levels = c("Mussels only", "Mussels and algae")) #specify factor levels in the order you want  (M)  (MA)

#to order the pCO2 levels in an increasing way
MSur.dat_mod2$pCO2 <- factor(MSur.dat_mod2$pCO2, levels = c("650 ppm", "450 ppm (ambient)")) #specify factor levels in the order you want

#to order the Temperature levels in an increasing way
MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient +2°C", "ambient +1°C", "ambient (9 - 12°C)")) #specify factor levels in the order you want

#to order the combined Treatment levels in an increasing way
MSur.dat_mod2$Treatment <- factor(MSur.dat_mod2$Treatment, levels = c("MA_Ta_Ca", "MA_Ta_C+", "MA_T+_Ca", "MA_T+_C+", "MA_T++_Ca", "MA_T++_C+", "M_Ta_Ca", "M_Ta_C+", "M_T+_Ca", "M_T+_C+", "M_T++_Ca", "M_T++_C+")) #specify factor levels in the order you want



FigPDP<-ggplot(MSur.dat_mod2, aes(x=Treatment, y=PD, colour=Temperature, fill=pCO2)) # colour needs to be character
FigPDP+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  geom_boxplot(lwd=0.8, width=0.7)+ #plots the data 
  scale_colour_manual(values = c("skyblue", "royalblue", "navy"))+ 
  scale_fill_manual(values = c("grey80", "white"))+  #, labels = expression("...", "...") 
  labs(x="Experimental treatments", y="Total mussel deaths [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_blank(), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = guide_legend(order = 2, title="Temperature"), fill = guide_legend(order = 1, title="pCO2"))
```
```{r}
#ggsave(path="Plots", filename="MusselTotalDeathsPercentage_Presentations.png", width=10, height=5)
```
Removing pCO2 as a separate factor, according to the statistical analysis
```{r}
#to order the Temperature levels in an increasing way
MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient (9 - 12°C)", "ambient +1°C", "ambient +2°C")) #specify factor levels in the order you want

FigPDPS<-ggplot(MSur.dat_mod2, aes(x=Temperature, y=PD, colour=Temperature)) # colour needs to be character
FigPDPS+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  geom_boxplot(lwd=0.8, width=0.7)+ #plots the data 
  scale_colour_manual(values = c("navy", "royalblue", "skyblue"))+ 
  #scale_fill_manual(values = c("grey80", "white"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Total mussel deaths [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_blank(), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = guide_legend(order = 2, title="Temperature"), fill = guide_legend(order = 1, title="pCO2"))
```
```{r}
#ggsave(path="Plots", filename="MusselTotalDeathsPercentage_Presentations_nopCO2.png", width=10, height=5)
```

```{r}
mean(PD[MSur.dat$Assemblage=="M"]) #31.85
mean(PD[MSur.dat$Assemblage=="KM"]) #24.26
mean(PD[MSur.dat$Temperature=="Ta"]) #22.64
mean(PD[MSur.dat$Temperature=="T++"]) #34.06
mean(PD[MSur.dat$Assemblage=="M"&MSur.dat$Temperature=="Ta"]) #23.59
mean(PD[MSur.dat$Assemblage=="KM"&MSur.dat$Temperature=="Ta"]) #21.69
mean(PD[MSur.dat$Assemblage=="M"&MSur.dat$Temperature=="T++"]) #40.18
mean(PD[MSur.dat$Assemblage=="KM"&MSur.dat$Temperature=="T++"]) #27.93

```



# 4. Analysis - does the amount of total survivors, or deaths respecitvely, differ between treatments?
## 4.1 multi-factorial ANOVA on percent deaths

aov() provides a wrapper to lm() for fitting linear models to balanced or unbalanced experimental designs.
The main difference from lm() is in the way print, summary and so on handle the fit: this is expressed in the traditional language of the analysis of variance rather than that of linear models.

first use lm / aov to fit a model, then use anova to analyse the result


aov() fits a model, so it produces regression coefficients, fitted values, residuals, etc; It produces an object of primary class "aov" but also a secondary class "lm". So, it is an augmentation of an "lm" object.
anova() is a generic function that returns a much more succinct type I (sequential) ANOVA table.

```{r}
table(MSur.dat$Treatment)
```


```{r}
Model1<-aov(MSur.dat$PD ~ MSur.dat$Assemblage*MSur.dat$Temperature*MSur.dat$pCO2) #Model1 multi-factorial
```
(assumptions checked in console or , all ok)
```{r}
#Model simplification (do you do model simplification with ANOVAs?)
summary(Model1) #main effects of assemblage and temperature significant (<0.01), rest insignificant
anova(Model1)

#remove insignificant 3-way interaction
Model1u1 <-update(Model1,~.-MSur.dat$Assemblage:MSur.dat$Temperature:MSur.dat$pCO2, data= MSur.dat)
anova(Model1, Model1u1) #removal ok if p>0.05, which is the case here
summary(Model1u1) #call summary to see which next parts can be removed from the model, starting with the longest interaction and, if there's more of the same type, the highest p-value

#sequentially remove all two-way interactions. Try them all individually first, before actually removing them
Model1u2a <-update(Model1u1,~.-MSur.dat$Temperature:MSur.dat$pCO2, data= MSur.dat)
anova(Model1u1, Model1u2a) #ok

Model1u2b <-update(Model1u1,~.-MSur.dat$Assemblage:MSur.dat$pCO2, data= MSur.dat)
anova(Model1u1, Model1u2b) #ok

Model1u2c <-update(Model1u1,~.-MSur.dat$Assemblage:MSur.dat$Temperature, data= MSur.dat)
anova(Model1u1, Model1u2c) #ok

Model1u3 <-update(Model1u1,~.-MSur.dat$Temperature:MSur.dat$pCO2, data= MSur.dat)
anova(Model1u1, Model1u3) #ok
summary(Model1u3)

Model1u4 <-update(Model1u3,~.-MSur.dat$Assemblage:MSur.dat$pCO2, data= MSur.dat)
anova(Model1u3, Model1u4) #ok
summary(Model1u4)

Model1u5 <-update(Model1u4,~.-MSur.dat$Assemblage:MSur.dat$Temperature, data= MSur.dat)
anova(Model1u4, Model1u5) #ok
summary(Model1u5)

Model1u6 <-update(Model1u5,~.-MSur.dat$pCO2, data= MSur.dat)
anova(Model1u5, Model1u6) #ok
summary(Model1u6)

#Model1u6 is the minimum adequate model
summary(Model1u6)
anova(Model1u6)

#Diagnostics for Model1u6
#homogeneity of variances
#graphical
plot(resid(Model1u6)~fitted(Model1u6))+
abline(h=0, lwd=2, col="black") #okish
#statistical
fligner.test(MSur.dat$PD ~ MSur.dat$Assemblage) #insignificant, i.e. so variances can be assumed to be homogeneous
fligner.test(MSur.dat$PD ~ MSur.dat$Temperature)
fligner.test(MSur.dat$PD ~ MSur.dat$pCO2)
fligner.test(MSur.dat$PD ~ interaction(MSur.dat$Assemblage, MSur.dat$Temperature, MSur.dat$pCO2))
leveneTest(MSur.dat$PD ~ MSur.dat$Assemblage*MSur.dat$Temperature*MSur.dat$pCO2) #requires library(car)
leveneTest(MSur.dat$PD ~ MSur.dat$Assemblage*MSur.dat$Temperature)

#normality of errors
#graphical
hist(resid(Model1u6)) #okish
#statistical (Shapiro-Wilk test works for 7 < n < 70)
shapiro.test(resid(Model1u6)) #insignificant, all good

#influential data points
plot(cooks.distance(Model1u6),type="h") #all <1, i.e. not problematic

#autocorrelation (requires library(car))
durbinWatsonTest(Model1u6) #borderline ok

AIC(Model1,Model1u6)
#Model1u6 is better than Model1 (value is smaller)
```
Model1
                                                       Df Sum Sq Mean Sq F value  Pr(>F)   
MSur.dat$Assemblage                                     1    864   864.3   8.403 0.00563 **
MSur.dat$Temperature                                    2   1313   656.5   6.382 0.00348 **
MSur.dat$pCO2                                           1     37    37.5   0.365 0.54885   
MSur.dat$Assemblage:MSur.dat$Temperature                2    276   138.0   1.342 0.27108   
MSur.dat$Assemblage:MSur.dat$pCO2                       1     62    62.3   0.606 0.44020   
MSur.dat$Temperature:MSur.dat$pCO2                      2      7     3.7   0.036 0.96455   
MSur.dat$Assemblage:MSur.dat$Temperature:MSur.dat$pCO2  2     74    37.0   0.360 0.69952   
Residuals                                              48   4937   102.9  


Model1u6
                     Df Sum Sq Mean Sq F value  Pr(>F)   
MSur.dat$Assemblage   1    864   864.3   8.972 0.00408 **
MSur.dat$Temperature  2   1313   656.5   6.815 0.00224 **
Residuals            56   5394    96.3    


### 4.1.2 Retrospective power analysis
```{r}
#final output
anova(Model1)
summary(Model1u6)
anova(Model1u6)
```

### 4.1.3 Retrospective power analysis
groups = number of factor levels
n = number of observations per sample
between.var = mean squares of the effect (sum sq)
within.var = mean squares of the residuals (mean sq)
sig.level = error rate (default is 0.05)
power = the power you want to achieve

NULL must be replaced by observed values. If you leave one of them as NULL,
this one will be calculated from the rest.

```{r}
power.anova.test(groups=12, n=5, between.var=864.29, within.var=96.3, sig.level=0.05, power=NULL) #high power to detect assemblage effect
power.anova.test(groups=12, n=5, between.var=656.46, within.var=96.3, sig.level=0.05, power=NULL) #high power to detect temperature effect
power.anova.test(groups=12, n=5, between.var=37.49, within.var=102.86, sig.level=0.05, power=NULL) #power of 0.79 to detect/rule out pCO2 effect
```


### 4.1.4 Find differences - Post-hoc comparisons
include Bonferroni Correction to account for increased Type I error probability due to high number of comparisons
Tukey's test
```{r}
TukeyHSD(Model1)
TukeyHSD(Model1u6)
```
KM statistically different from M, and Ta from T++

##4.2 mean-error plots
```{r}
#for better facet labeling I need to rename the levels of Assemblage:
MSur.dat_mod2 <- MSur.dat %>%    #requires tidyverse
  # Rename levels of Assemblage
  mutate(Assemblage = dplyr::recode(Assemblage, "KM" = "Mussels and algae", "M" = "Mussels only"), # (MA)  (M)
         pCO2 = dplyr::recode(pCO2, "C+" = "650 ppm", "Ca" = "450 ppm \n(ambient)"),
         Temperature = dplyr::recode(Temperature, "T++" = "+2°C", "T+" = "+1°C", "Ta" = "ambient \n(9 - 12°C)"),
         Treatment = dplyr::recode(Treatment, "KM_Ta_Ca"="MA_Ta_Ca", "KM_Ta_C+"="MA_Ta_C+", "KM_T+_Ca"="MA_T+_Ca", "KM_T+_C+"="MA_T+_C+", "KM_T++_Ca"="MA_T++_Ca", "KM_T++_C+"="MA_T++_C+"))

#to place Mussels only first in the order of the Assemblage levels
MSur.dat_mod2$Assemblage <- factor(MSur.dat_mod2$Assemblage, levels = c("Mussels only", "Mussels and algae")) #specify factor levels in the order you want  (M)  (MA)

#to order the pCO2 levels in an increasing way
MSur.dat_mod2$pCO2 <- factor(MSur.dat_mod2$pCO2, levels = c("450 ppm \n(ambient)", "650 ppm")) #specify factor levels in the order you want

#to order the Temperature levels in an increasing way
#MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient +2°C", "ambient +1°C", "ambient (9 - 12°C)")) #specify factor levels in the order you want

#to order the combined Treatment levels in an increasing way
MSur.dat_mod2$Treatment <- factor(MSur.dat_mod2$Treatment, levels = c("MA_Ta_Ca", "MA_Ta_C+", "MA_T+_Ca", "MA_T+_C+", "MA_T++_Ca", "MA_T++_C+", "M_Ta_Ca", "M_Ta_C+", "M_T+_Ca", "M_T+_C+", "M_T++_Ca", "M_T++_C+")) #specify factor levels in the order you want


```

```{r all data}
#to order the Temperature levels in an increasing way
MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient \n(9 - 12°C)", "+1°C", "+2°C")) #specify factor levels in the order you want

FigPD<-ggplot(MSur.dat_mod2, aes(x=Temperature, y=PD, colour=Temperature, fill=pCO2)) # colour needs to be character
FigPD+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, linewidth=1.3, position="dodge")+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+
  scale_fill_manual(values = c("white","grey80"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Mussel deaths [%]")+ #labeling of axes
  scale_y_continuous(breaks=seq(0,50,by=10))+
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_line(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=12, colour="black"), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12, colour="black"),
        axis.title=element_text(size=12),#size of the labels on the axes 
        legend.position = "bottom", #c(0.9, 0.9),
        legend.text = element_text(size=12), #
        legend.title = element_text(size=12),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = guide_legend(order = 1, title=bquote("pCO"[2]~"level"), override.aes=list(colour="black", linewidth=1))) #guide_legend(order = 2, title="Temperature")+ guide_legend(order = 1, title="Temperature")
  #annotate("text", x=3, y=59, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2023-01-23_Mortality_mean-errorSE_all-data.png", width=5, height=4)
```

```{r main effect temperature}
FigPDT<-ggplot(MSur.dat_mod2, aes(x=Temperature, y=PD, colour=Temperature, fill=Temperature)) # colour needs to be character
FigPDT+
  #facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, linewidth=1.3, position="dodge")+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  scale_fill_manual(values = c("skyblue", "royalblue", "blue3"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Mussel deaths [%]")+ #labeling of axes
  scale_y_continuous(breaks=seq(0,40,by=10))+
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_line(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=12, colour="black"), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12, colour="black"),
        axis.title=element_text(size=12),#size of the labels on the axes 
        legend.position = "bottom", #c(0.9, 0.9),
        legend.text = element_text(size=12), #
        legend.title = element_text(size=12),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = "none")+ #guide_legend(order = 2, title="Temperature")+ guide_legend(order = 1, title="Temperature") guide_legend(order = 1, title=bquote("pCO"[2]~"level"), override.aes=list(colour="black", linewidth=1))
  annotate("text", x=3, y=42, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2023-01-23_Mortality_mean-errorSE_Temperature-main-effect.png", width=2.5, height=2)
```

```{r main effect species blank}
MSur.dat_mod3 <- MSur.dat %>%    #requires tidyverse
  # Rename levels of Assemblage
  mutate(Assemblage = dplyr::recode(Assemblage, "KM" = "Mussels \nand algae", "M" = "Mussels \nonly"))
#to place Mussels only first in the order of the Assemblage levels
MSur.dat_mod3$Assemblage <- factor(MSur.dat_mod3$Assemblage, levels = c("Mussels \nonly", "Mussels \nand algae")) #specify factor levels in the order you want  (M)  (MA)
         
FigPDSb<-ggplot(MSur.dat_mod3, aes(x=Assemblage, y=PD)) # colour needs to be character
FigPDSb+
  #facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, linewidth=1, position="dodge", colour="black", fill="white")+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  #scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  #scale_fill_manual(values = c("skyblue", "royalblue", "blue3"))+  #, labels = expression("...", "...") 
  labs(x="Presence of algae", y="Mussel deaths [%]")+ #labeling of axes
  scale_y_continuous(breaks=seq(0,40,by=10))+
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_line(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=12, colour="black"), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12, colour="black"),
        axis.title.y=element_text(size=12),#size of the labels on the axes 
        axis.title.x=element_blank(),
        legend.position = "bottom", #c(0.9, 0.9),
        legend.text = element_text(size=12), #
        legend.title = element_text(size=12),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = "none")+ #guide_legend(order = 2, title="Temperature")+ guide_legend(order = 1, title="Temperature") guide_legend(order = 1, title=bquote("pCO"[2]~"level"), override.aes=list(colour="black", linewidth=1))
  annotate("text", x=3, y=42, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2023-01-23_Mortality_mean-errorSE_Species-main-effect.png", width=2.5, height=2)
```


```{r both significant main effects}
#to order the Temperature levels in an increasing way
MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient (9 - 12°C)", "ambient +1°C", "ambient +2°C")) #specify factor levels in the order you want

FigPDme<-ggplot(MSur.dat_mod2, aes(x=Temperature, y=PD, colour=Temperature, fill=Temperature)) # colour needs to be character
FigPDme+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", linewidth=0.7)+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  
  scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  scale_fill_manual(values = c("skyblue", "royalblue", "blue3"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Mussel deaths [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_blank(), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = guide_legend(order = 1, title="Temperature"))+ #guide_legend(order = 2, title="Temperature")+
  annotate("text", x=3, y=46, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2022-11-23_Survival_mean-errorSE_flipped-colours_removed-abbreviations.png", width=7, height=4)
```


```{r both significant main effects}
#to order the Temperature levels in an increasing way
MSur.dat_mod2$Temperature <- factor(MSur.dat_mod2$Temperature, levels = c("ambient (9 - 12°C)", "ambient +1°C", "ambient +2°C")) #specify factor levels in the order you want

FigPDme<-ggplot(MSur.dat_mod2, aes(x=Temperature, y=PD, colour=Temperature, fill=Temperature)) # colour needs to be character
FigPDme+
  facet_wrap(.~Assemblage, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", linewidth=0.7)+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  
  scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  scale_fill_manual(values = c("skyblue", "royalblue", "blue3"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Mussel deaths [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_blank(), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = guide_legend(order = 1, title="Temperature"))+ #guide_legend(order = 2, title="Temperature")+
  annotate("text", x=3, y=46, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2022-11-23_Survival_mean-errorSE_flipped-colours_removed-abbreviations.png", width=7, height=4)
```