---
title: "Mussel condition index 10-12/2020"
author: "Katrin Schertenleib"
date: "`r format(Sys.Date(), '%d/%m/%y')`"
output: github_notebook
---
file first created on Sun Sep 18 15:32:51 2022

aim: determine and plot the mussel condition index

# 1. Housekeeping
libraries
```{r setup} 
#library(installr) #to always use the latest version
    #updateR()
#library(readxl)   #to import data from excel
library(yaml)     #for advanced Rmarkdown layout options, e.g. using code in the header (allows to print current date in Markdown header)
library(tidyverse)#includes ggplot, dplyr and plyr
#library(lubridate) #to run parse_date_time
#library(nortest) #to run Lilliefors corrected Kolmogorov-Smirnov normality tests
library(car) #to run Anova() with different type sums of squares
library(Hmisc) #to calculate stats for mean error plots in ggplot
```
note: If you have a code block named 'setup' like```{r setup} foo() ``` then every time you restart RStudio and execute any code in the middle of your markdown document, this block will be automatically run once before, i.e. you libraries will be loaded first.
foo() is a placeholder variable in coding

check work environment (especially, if you're not working in R projects)
```{r}
getwd() #when working in an R project, no need to change this manually. setwd() allows you to set it manually
#dir() #lists what is in the work directory
```



# 2. Reading in the data
```{r}
MWWout.dat <- read_csv("2020-12-17to18_MesocosmExperiment_MusselsOUT.csv") #mussel data at the end of the experiment
```

# 3. Data preparation and exploration
## 3.1 check and clean
check, if data was read in correctly
```{r}
spec(MWWout.dat)
str(MWWout.dat) #displays internal structure of the objects in the data frame
names(MWWout.dat) #lists all variable headers in the data frame
nrow(MWWout.dat) #gives you number of rows in the data frame

ncol(MWWout.dat) #gives you number of columns in the data frame
head(MWWout.dat) #good way to check the first 6 rows, i.e. see if not only the headers but also the data looks ok
tail(MWWout.dat) #looks at the end of the data set
```

reduce data frame; separate the combined treatment information into additional columns for each factor
```{r}
MWWout.dat <- MWWout.dat[,c(4,6,22,23,28,30,31)]
MWWout.dat <- separate(MWWout.dat, treatment, sep = "_", into = c("SpeciesLevel", "TempLevel", "pCO2Level"), remove=FALSE)
```

## 3.2 Calculations
[g]
```{r}
MWWout.dat$DWFlesh <- MWWout.dat$DWSubsampleFleshOnALf - MWWout.dat$Alf
MWWout.dat$DWShell <- MWWout.dat$DWSubsampleShellOnALs - MWWout.dat$Als
MWWout.dat$CI <- (MWWout.dat$DWFlesh/MWWout.dat$DWShell)*100
```
```{r}
MWWout.dat <- subset(MWWout.dat, MWWout.dat$SpeciesLevel=="KM" | MWWout.dat$SpeciesLevel=="M")
```

```{r}
summary(MWWout.dat)
```


```{r}
#write.csv(MWWout.dat,"2022-09-18_2020-12-17_MesocosmExperiment_MusselConditionIndex.csv", row.names = TRUE) #export 
```

# 4. Exploratory Plotting
```{r}
#to order the temperature levels in an increasing way in the plot (lowest on bottom, highest on top)
MWWout.dat$TempLevel <- factor(MWWout.dat$TempLevel, levels = c("Ta", "T+", "T++")) #specify factor levels in the order you want
MWWout.dat$pCO2Level <- factor(MWWout.dat$pCO2Level, levels = c("Ca", "C+")) #specify factor levels in the order you want

#reduce to M and KM treatments only
MWWoutMKM.dat <- MWWout.dat[MWWout.dat$SpeciesLevel=="M" | MWWout.dat$SpeciesLevel=="KM",]

#for better facet labeling I need to rename the levels of Assemblage:
MWWoutMKM.dat_mod <- MWWoutMKM.dat %>%    #requires tidyverse
  # Rename levels of Assemblage
  mutate(SpeciesLevel = dplyr::recode(SpeciesLevel, "B" = "Blank (B)", "K" = "Kelp (K)", "KM" = "Mussels and kelp (KM)", "M" = "Mussels only (M)"))
```

```{r}
names(MWWoutMKM.dat)
```
```{r Mussel Condition index}
FigureMCI <- ggplot(MWWoutMKM.dat_mod, aes(x=TempLevel, y=CI, colour=TempLevel, fill=pCO2Level))  
FigureMCI+
  geom_boxplot()+ #, lwd=0.7 width=0.6, position=position_dodge(0.5)
  facet_wrap(.~SpeciesLevel)+
  labs(y="Mussel condition index [%]", x="Temperature Level")+ #labelling of axes (can be removed by setting to FALSE)
  scale_colour_manual(values = c("navy", "royalblue", "skyblue"))+ #manually select colours of "fill"
  scale_fill_manual(values=c("white","grey80"))+
  #scale_x_discrete(labels=c("5"= "Dec 08\n09:30", "6"= "Dec 08\n13:30", "7"= "Dec 08\n17:30", "8"= "Dec 09"))+
  #scale_y_continuous(breaks=seq(7,16,by=1))+
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black", size= 0.5), #colour of the x- axis
        axis.line.y=element_line(colour="black", size= 0.5), #colour of the y- axis
        axis.ticks.x = element_line(), #x axis tick; to include and modify width use element_line(size = 0.8)
        axis.title.x=element_text(size=9),
        panel.grid=element_blank(), #no grid lines
        axis.text=element_text(size=9), #size of the numbers along the axes
        #axis.text.x=element_blank(), #x axis tick mark font specifications; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=9),
        axis.title=element_text(size=9), #size of the labels on the axes
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=9), #
        legend.title = element_text(size=9),
        #legend.position = c(0.92, 0.85),
        #legend.key=element_rect(fill="white"),
        #legend.background = element_rect(colour = "black")
        #strip.text = element_text(size = 9), #font size of facet labels
        #strip.background = element_blank() #removes facet label boxes
        #strip.background = element_rect(colour= "black", fill="grey97"))
        )+
  guides(fill = guide_legend(order = 2, title="pCO2 Level"), colour = guide_legend(order=1, title="Temperature \nLevel"))
```
```{r}
ggsave(path="Plots", filename="2022-09-18_MusselConditionIndex.png", width=9, height=3.5) #legend on figure
#ggsave(path="Plots", filename="SnapshotTemperatureMeasurements3.png", width=11, height=4)#legend right of figure
```

# 5. Statistical Analyses
aov() provides a wrapper to lm() for fitting linear models to balanced or unbalanced experimental designs.
The main difference from lm() is in the way print, summary and so on handle the fit: this is expressed in the traditional language of the analysis of variance rather than that of linear models.

first use lm / aov to fit a model, then use anova to assess if removing a predictor improves the fit of the model

aov() fits a model, so it produces regression coefficients, fitted values, residuals, etc; It produces an object of primary class "aov" but also a secondary class "lm". So, it is an augmentation of an "lm" object.
anova() is a generic function that returns a much more succinct type I (sequential) ANOVA table.

If you want to run Tukey’s HSD tests later, consider that it requires an aov object on which it runs its procedure

"anova()" is a function in base R and uses type I Sums of Squares as default, 
"Anova()" is from the package 'car' and uses type II or III sos

In ANOVA, you can use different types of sums of squares (https://books.google.ch/books?id=wd2K2zC3swIC&lpg=PP1&hl=de&pg=PA475#v=onepage&q&f=false)
Type I: sequential
Type II: most powerful for main effects and no (!) predicted interactions
Type III: when interactions are present, the main effects associated with that interaction will still be meaningful; requires orthogonal contrasts; can be used for inbalanced design 
Anova(Model1, type=3) (note capital A!!!) needs the following general specification: options(contrasts=c("contr.sum", "contr.poly"))

```{r}
summary(MWWout.dat)
names(MWWout.dat)
```

```{r}
hist(MWWout.dat$CI)
hist(sqrt(MWWout.dat$CI)) 
hist(log(MWWout.dat$CI))
```
r for raw (untransformed) data
```{r}
options(contrasts=c("contr.sum", "contr.poly"))
Model1r<-aov(MWWout.dat$CI ~ MWWout.dat$SpeciesLevel*MWWout.dat$TempLevel*MWWout.dat$pCO2Level) #Model1 multi-factorial; add +0.1 so that roots and logs make sense
```


```{r}
#Model simplification (do you do model simplification with ANOVAs?)
summary(Model1r) #main effects of species level significant (0.0363), rest insignificant
Anova(Model1r)


#Diagnostics for Model1r
#homogeneity of variances
#graphical
plot(resid(Model1r)~fitted(Model1r))+
abline(h=0, lwd=2, col="black") #ok
#statistical
leveneTest(MWWout.dat$CI ~ MWWout.dat$SpeciesLevel*MWWout.dat$TempLevel*MWWout.dat$pCO2Level)
leveneTest(MWWout.dat$CI ~ MWWout.dat$SpeciesLevel) #requires library(car)
leveneTest(resid(Model1r) ~MWWout.dat$SpeciesLevel)

#normality of errors
#graphical
hist(resid(Model1r)) #ok
#statistical (Shapiro-Wilk test works for 7 < n < 70)
shapiro.test(resid(Model1r)) #insignificant, i.e. data normally distributed

#influential data points
plot(cooks.distance(Model1r),type="h") #all <1, i.e. not problematic

#autocorrelation (requires library(car))
durbinWatsonTest(Model1r) #DW statistic ranges from 0 to 4, with 2 indicating no autocorrelation. <2 is positively autocorrelated, >2 is negatively autocorrelated. Values in the range of 1.5 - 2.5 are considered "normal"
#hm. slight positive autocorrelation? might increase type I errors
```
full model
                                                                  Df Sum Sq Mean Sq F value   Pr(>F)    
MWWout.dat$SpeciesLevel                                            1 111.48  111.48  18.437 8.49e-05 ***
MWWout.dat$TempLevel                                               2   1.04    0.52   0.086    0.917    
MWWout.dat$pCO2Level                                               1   0.56    0.56   0.093    0.762    
MWWout.dat$SpeciesLevel:MWWout.dat$TempLevel                       2   2.97    1.49   0.246    0.783    
MWWout.dat$SpeciesLevel:MWWout.dat$pCO2Level                       1   9.09    9.09   1.503    0.226    
MWWout.dat$TempLevel:MWWout.dat$pCO2Level                          2   1.10    0.55   0.091    0.913    
MWWout.dat$SpeciesLevel:MWWout.dat$TempLevel:MWWout.dat$pCO2Level  2  27.05   13.52   2.237    0.118    
Residuals                                                         48 290.22    6.05                     
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

reduced model
                        Df Sum Sq Mean Sq F value  Pr(>F)    
MWWout.dat$SpeciesLevel  1  111.5  111.48   19.47 4.5e-05 ***
Residuals               58  332.0    5.72      



## 5.3 Retrospective power analysis
groups = number of factor levels
n = number of observations per sample
between.var = mean squares of the effect
within.var = mean squares of the residuals
sig.level = error rate (default is 0.05)
power = the power you want to achieve

NULL must be replaced by observed values. If you leave one of them as NULL,
this one will be calculated from the rest.

17 h main effect temperature
```{r}
power.anova.test(groups=12, n=5, between.var=111.48, within.var=5.72, sig.level=0.05, power=NULL) #high power (1)
```

## 5.4 Find differences - Post-hoc comparisons
include Bonferroni Correction to account for increased Type I error probability due to high number of comparisons
Tukey's test
NOT NECESSARY HERE BECAUSE ONLY 2 LEVELS
```{r}
TukeyHSD(Model1r)
```

# 6. Plotting significant effects
```{r}
#for better facet labeling I need to rename the levels of SpeciesLevel:
MWWout.dat_mod2 <- MWWout.dat %>%    #requires tidyverse
  # Rename levels of SpeciesLevel
  mutate(SpeciesLevel = dplyr::recode(SpeciesLevel, "KM" = "Mussels and algae", "M" = "Mussels only"),
         pCO2Level = dplyr::recode(pCO2Level, "C+" = "650 ppm", "Ca" = "450 ppm (ambient)"),
         TempLevel = dplyr::recode(TempLevel, "T++" = "+2°C", "T+" = "+1°C", "Ta" = "ambient \n(9 - 12°C)"),
         treatment = dplyr::recode(treatment, "KM_Ta_Ca"="MA_Ta_Ca", "KM_Ta_C+"="MA_Ta_C+", "KM_T+_Ca"="MA_T+_Ca", "KM_T+_C+"="MA_T+_C+", "KM_T++_Ca"="MA_T++_Ca", "KM_T++_C+"="MA_T++_C+"))

#to place Mussels only first in the order of the SpeciesLevel levels
MWWout.dat_mod2$SpeciesLevel <- factor(MWWout.dat_mod2$SpeciesLevel, levels = c("Mussels only", "Mussels and algae")) #specify factor levels in the order you want

#to order the pCO2 levels in an increasing way
MWWout.dat_mod2$pCO2Level <- factor(MWWout.dat_mod2$pCO2Level, levels = c("450 ppm (ambient)", "650 ppm")) #specify factor levels in the order you want

#to order the Temperature levels in an increasing way
MWWout.dat_mod2$TempLevel <- factor(MWWout.dat_mod2$TempLevel, levels = c("ambient \n(9 - 12°C)", "+1°C", "+2°C")) #specify factor levels in the order you want
```

```{r CI all data}
FigCImeall <-ggplot(MWWout.dat_mod2, aes(x=TempLevel, y=CI, colour=TempLevel, fill=pCO2Level)) # colour needs to be character , colour=TempLevel, fill=TempLevel , fill=SpeciesLevel
FigCImeall+
  facet_wrap(.~SpeciesLevel, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, linewidth=1.3, position="dodge")+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc) linewidth=0.7,size=10, , colour="black" , fill= "grey70" , fill=c("grey75", "grey50")
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.8), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data 
  scale_y_continuous(breaks=seq(0,24,by=5))+
  scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  scale_fill_manual(values = c("white", "grey80"))+  #, labels = expression("...", "...") 
  labs(x="Temperature level", y="Mussel condition index [%]")+ #labeling of axes
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_line(size=0.8), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=12, colour="black"), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12, colour="black"),
        axis.title=element_text(size=12),#size of the labels on the axes 
        #axis.title.y=element_text(size=12),#size of the labels on the axes
        #axis.title.x=element_blank(),
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = "none")+ #guide_legend(order = 2, title="Temperature")+
  annotate("text", x=1, y=29, label= " ") #guide_legend(order = 1, title=bquote("pCO"[2]~"level"), override.aes=list(colour="black", linewidth=1))
```

```{r}
ggsave(path="Plots", filename="2023-01-25_MCI_mean-errorSE_all-data.png", width=7, height=4)#width=7 or 5
```

```{r mussel daily wet weigt increase only species main effect}
FigCImeS<-ggplot(MWWout.dat_mod2, aes(x=SpeciesLevel, y=CI)) # colour needs to be character , colour=TempLevel, fill=TempLevel , fill=SpeciesLevel
FigCImeS+
  #facet_wrap(.~SpeciesLevel, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, linewidth=1, position="dodge", colour="black", fill="white")+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc)
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  #scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  #scale_fill_manual(values = c("grey70", "grey90"))+  #, labels = expression("...", "...") 
  labs(x="none", y="none")+ #labeling of axes
  scale_y_continuous(breaks=seq(0,20,by=5))+
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_blank(),#text(size=12), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=16),
        axis.title=element_blank(),#text(size=12),#size of the labels on the axes 
        #axis.title.y=element_text(size=12),#size of the labels on the axes
       # axis.title.x=element_blank(),
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=16), #
        #legend.title = element_text(),
        strip.text = element_text(size = 16), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  #guides(colour = "none", fill = guide_legend(order = 1, title="Temperature"))+ #guide_legend(order = 2, title="Temperature")+
  annotate("text", x=1, y=22, label= " ") 
```
```{r}
ggsave(path="Plots", filename="2023-01-25_MCI_mean-errorSE_species-main-effect.png", width=3, height=2)#width=7 or 5
```


```{r CI main effect old}
FigCIme<-ggplot(MWWout.dat_mod2, aes(x=SpeciesLevel, y=CI)) # colour needs to be character , colour=TempLevel, fill=TempLevel , fill=SpeciesLevel
FigCIme+
  #facet_wrap(.~SpeciesLevel, scales="free_x")+ #, strip.position = "bottom"
  stat_summary(fun=mean, geom="bar", width=0.8, fill=c("grey75", "grey50"))+ # Layer 2: Plot the means as bars , colour="white"  NOTE: requires library(Hmisc) linewidth=0.7,size=10, , colour="black" , fill= "grey70"
  stat_summary(fun.data=mean_se, geom="errorbar", position=position_dodge(width=0.9), width=0.0, colour="black", linetype=1, linewidth=0.8)+# Layer 3: Add CIs as error bars using mean_cl_normal as fun.data
  
  #scale_colour_manual(values = c("skyblue", "royalblue", "blue3"))+ 
  #scale_fill_manual(values = c("grey70", "grey90"))+  #, labels = expression("...", "...") 
  labs(y="Mussel condition index [%]")+ #labeling of axes x="none", 
  theme(panel.background=element_blank(), #no background
        axis.line.x=element_line(colour="black"), #colour of the x- axis; size= 0.8
        axis.line.y=element_line(colour="black"), #colour of the y- axis; size= 0.8
        axis.ticks.x = element_blank(), #x axis tick; to include and modify width use element_line(size = 0.8)
        #axis.ticks.length=unit(),#.25, "cm"
        panel.grid=element_blank(), #no grid lines
        axis.text.x=element_text(size=12), #for size of the numbers along the axes; include labels via element_text(size=10, angle=90, hjust=0, vjust=0.25); hjust and vjust are defined between 0(left-justified) and 1(right justified)
        axis.text.y=element_text(size=12),
        axis.title=element_text(size=12),#size of the labels on the axes 
        axis.title.y=element_text(size=12),#size of the labels on the axes
        axis.title.x=element_blank(),
        #legend.position = c(0.9, 0.9),
        legend.text = element_text(size=12), #
        #legend.title = element_text(),
        strip.text = element_text(size = 12), #font size of facet labels
        strip.background = element_blank() #removes facet label boxes
        )+
  guides(colour = "none", fill = guide_legend(order = 1, title="Temperature"))+ #guide_legend(order = 2, title="Temperature")+
  annotate("text", x=1, y=19, label= " ") 
```

```{r}
ggsave(path="Plots", filename="2022-11-23_MConditionIndex_mean-errorSE2.png", width=3.5, height=4)#width=7 or 5
```

```{r}
tapply(MWWout.dat$CI, MWWout.dat$SpeciesLevel, mean)
```


